---
author: Analysis Documentation
date: June 2026
output: html_document
title: Christchurch Airbnb Listings Analysis --- Data Preparation and
  Outputs
---

# 1. Overview

This R Markdown document documents the data preparation, merging,
summary inspection, CSV export, and visualisations produced from the
Christchurch Airbnb listings datasets.

The analysis uses the `tidyverse` package and combines monthly
Christchurch listing datasets covering **October 2025 to June 2026**.

`{r setup, message=FALSE, warning=FALSE} library(tidyverse)`

# 2. Part One --- Filter and Prepare the June 2026 Dataset

## 2.1 Christchurch filtering

The June 2026 listings dataset is filtered so that only records
belonging to **Christchurch City** are retained.

`{r filter-christchurch} chcListingsJun2026 <- listingsJun2026 %>%   filter(neighbourhood_group == "Christchurch City")`

### Output

**Object created:** `chcListingsJun2026`

This dataframe contains the June 2026 listings for Christchurch City
only.

## 2.2 Add month and year variables

Two variables are added to identify the month and year represented by
the dataset.

\`\`\`{r add-month-year} chcListingsJun2026 \<- chcListingsJun2026 \|\>
mutate(data_month = "June")

chcListingsJun2026 \<- chcListingsJun2026 \|\> mutate(data_year =
"2026")


    ### Output

    The resulting dataframe contains:

    - `data_month` — `"June"`
    - `data_year` — `"2026"`

    These variables provide explicit temporal information for the June 2026 observations.

    # 3. Part Two — Merge Monthly Datasets

    The Christchurch datasets from October 2025 through June 2026 are combined into a single dataframe using `rbind()`.

    ```{r merge-datasets}
    chcListingsTotal <- rbind(
      chcListingsOct2025,
      chcListingsNov2025,
      chcListingsDec2025,
      chcListingsJan2026,
      chcListingsFeb2026,
      chcListingsMar2026,
      chcListingsApr2026,
      chcListingsMay2026,
      chcListingsJun2026
    )

## Output

**Object created:** `chcListingsTotal`

This dataframe represents the combined Christchurch listing data for:

-   October 2025
-   November 2025
-   December 2025
-   January 2026
-   February 2026
-   March 2026
-   April 2026
-   May 2026
-   June 2026

The purpose of this merged dataset is to provide one dataset for
subsequent Christchurch-wide analysis across the nine-month period.

# 4. Part Three --- Export the Merged Dataset

The combined dataframe is exported as a CSV file.

`{r export-csv} write.csv(chcListingsTotal, "chcListingsTotal.csv")`

## Output

**File produced:** `chcListingsTotal.csv`

The CSV contains the records and variables contained in
`chcListingsTotal`.

This provides a reusable external copy of the processed Christchurch
dataset.

# 5. Part Four --- Inspect Room-Type Categories

The unique room types in the combined dataset are identified.

\`\`\`{r room-categories} categories \<-
unique(chcListingsTotal\$room_type) numberOfCategories \<-
length(categories)

print(categories) print(numberOfCategories)


    ## Outputs

    ### `categories`

    A vector containing the unique values of the `room_type` variable.

    ### `numberOfCategories`

    The number of unique room-type categories.

    The values are printed to the R console for inspection.

    # 6. Part Five — Inspect Character Variables

    The code loops through every column in `chcListingsTotal` and identifies columns whose class is `character`.

    ```{r character-columns}
    for (i in colnames(chcListingsTotal)) {
      if (class(chcListingsTotal[[i]]) == "character") {
        print(chcListingsTotal[[i]])
        print(length(unique(chcListingsTotal[[i]])))
      }
    }

## Output

For every character column, the R console displays:

1.  The values contained in the column.
2.  The number of unique values in that column.

This provides a basic inspection of categorical/text variables in the
dataset.

# 7. Part Six --- Graph 1: Histogram of Listing Prices

A filtered version of the combined dataset is created before producing
the price histogram.

`{r price-filter} chcListingsTotalFilt <- chcListingsTotal %>%   filter(price < 3000)`

## Filtering criterion

Only listings with:

**Price \< NZD 3,000**

are included in the histogram.

Importantly, this filtering does **not** modify `chcListingsTotal`. It
creates a separate dataframe called `chcListingsTotalFilt`.

## Histogram

`{r price-histogram} hist(   chcListingsTotalFilt$price,   breaks = 50,   xlab = "Price of Listing (nzd)",   main = "Histogram of listing prices in Christchurch" )`

## Output

**Graph:** Histogram of Christchurch listing prices.

-   **X-axis:** Price of listing in NZD
-   **Y-axis:** Frequency of listings
-   **Number of bins:** 50
-   **Price restriction:** Listings below NZD 3,000

### Purpose

The histogram is used to examine the distribution of listing prices in
Christchurch while limiting the influence of extremely high-priced
listings.

# 8. Part Seven --- Graph 2: Days Since Last Review

The number of days since each listing's last review is calculated.

The reference/end date is:

**21 June 2026**

\`\`\`{r review-date-calculation} date_diffs \<- c()

end_date \<- as.POSIXlt("2026-06-21")

for (given_date in chcListingsTotal\$last_review) { if
(!is.na(given_date) & given_date != "") { start_date \<-
as.POSIXct(given_date) time_diff \<- difftime( end_date, start_date,
units ="days" ) numeric_days \<- as.numeric(time_diff) date_diffs \<-
c(date_diffs, numeric_days) } }


    ## Output

    **Object created:** `date_diffs`

    This numeric vector contains the calculated number of days between each valid `last_review` date and 21 June 2026.

    Listings with missing (`NA`) or empty `last_review` values are excluded from this calculation.

    ## Histogram

    ```{r review-histogram}
    hist(
      date_diffs,
      xlab = "Days since last review",
      main = "Histogram of days since last review"
    )

## Output

**Graph:** Histogram of days since the last review.

-   **X-axis:** Days since last review
-   **Y-axis:** Frequency of listings
-   **Reference date:** 21 June 2026

### Purpose

This visualisation shows how recently Christchurch listings received
reviews and the distribution of time elapsed since their most recent
review.

# 9. Part Eight --- Graph 3: Top 10% of Listings by Number of Reviews

The listings with the highest number of reviews are selected.

`{r top-reviews} chcListingsTop10Per <- chcListingsTotal %>%   slice_max(     order_by = number_of_reviews,     prop = 0.1   )`

## Output

**Object created:** `chcListingsTop10Per`

This dataframe contains approximately the **top 10% of listings based on
`number_of_reviews`**.

## Histogram

`{r review-count-histogram} hist(   chcListingsTop10Per$number_of_reviews,   xlab = "Number of Reviews",   main = "Histogram of 90th percentile of total reviews in a listing" )`

## Output

**Graph:** Histogram of the number of reviews among the highest-reviewed
listings.

-   **X-axis:** Number of reviews
-   **Y-axis:** Frequency of listings
-   **Selection:** Approximately top 10% of listings by number of
    reviews

### Terminology note

The code selects the **top 10% of listings**, rather than calculating
the 90th percentile value itself. Therefore, a more technically precise
description is:

> **Histogram of number of reviews for the top 10% of listings.**

# 10. Summary of Objects and Files Produced

  ------------------------------------------------------------------------
  Output                   Type                    Description
  ------------------------ ----------------------- -----------------------
  `chcListingsJun2026`     Dataframe               Christchurch City
                                                   listings for June 2026

  `chcListingsTotal`       Dataframe               Combined Christchurch
                                                   listings from October
                                                   2025 to June 2026

  `chcListingsTotal.csv`   CSV file                Exported combined
                                                   Christchurch dataset

  `categories`             Vector                  Unique room-type
                                                   categories

  `numberOfCategories`     Numeric value           Number of unique
                                                   room-type categories

  Character-column console Console output          Values and unique-value
  output                                           counts for character
                                                   columns

  `chcListingsTotalFilt`   Dataframe               Listings with price
                                                   below NZD 3,000

  `date_diffs`             Numeric vector          Days since last review
                                                   relative to 21 June
                                                   2026

  `chcListingsTop10Per`    Dataframe               Approximately top 10%
                                                   of listings by number
                                                   of reviews

  Price histogram          Graph                   Distribution of listing
                                                   prices

  Review-age histogram     Graph                   Distribution of days
                                                   since last review

  Top-review histogram     Graph                   Distribution of reviews
                                                   among the top 10% of
                                                   listings
  ------------------------------------------------------------------------

# 11. Overall Workflow

The complete analysis workflow can be summarised as:

**Filter Christchurch → Add temporal variables → Merge monthly datasets
→ Export CSV → Inspect categorical variables → Filter extreme prices →
Produce price histogram → Calculate review age → Produce review-age
histogram → Select top 10% by reviews → Produce review-count histogram**

# 12. Key Analytical Outputs

The code therefore produces three primary visualisations:

1.  **Listing Price Distribution**\
    Examines the distribution of Christchurch listing prices below NZD
    3,000.

2.  **Days Since Last Review Distribution**\
    Examines how recently Christchurch listings received their most
    recent review as of 21 June 2026.

3.  **Review Count Distribution for Top 10% of Listings**\
    Examines the number of reviews among the listings with the highest
    review counts.

Together, these outputs provide an initial descriptive analysis of
**listing prices, review activity, and highly reviewed listings** across
Christchurch.
